The `zoon` R package for reproducible and shareable species distribution modelling

### 2016-05-20 ms.Rmd

---
title: "The `zoon` R package for reproducible and shareable species distribution modelling"
author: "Nick Golding, Tom August, Tim C.D. Lucas, Emiel Van Loon & Greg McInerny"
date: "21 March 2016"
bibliography: zoon_app.bib
csl: mee.csl
output:
  pdf_document:
    fig_caption: yes
    includes:  
      in_header: latex_preamble.tex
---

@@

#plot the saved image in the first panel
library(raster)
r <- brick('figs/workflow_basic.png')
par(mar = c(1, 1, 2, 1))
plotRGB(r, maxpixels = Inf)

@@

```{r fengworkflow, eval = TRUE, dpi = 300, fig.show = 'hide', results = 'hide', message = FALSE}
Feng_Papes <- 
  workflow(occurrence = SpOcc('Dasypus novemcinctus',
                              extent = c(-130, -20, -60, 60)),
            covariate = Bioclim(extent = c(-130, -20, -60, 60),
                                layers = c(1:4, 6, 9, 10, 12, 15)),
              process = Chain(Clean, 
                              MESSMask,  
                              Background(n = 10000, bias = 200), 
                              Crossvalidate(k = 5)),
                model = MaxEnt,
               output = Chain(PerformanceMeasures, PrintMap))

@@

![Map of the presence data (red), background data (black) and the MaxEnt relative probability of occurrence (darker blue showing higher probability of occurrence). White areas are masked due to being marine in the marine environment or due to the MESS mask.\label{fig:fengmap}](./figs/fengworkflow-1.png)

@@

```{r fengChangeWorkflow, eval = TRUE, fig.show = 'hide'}
Feng_Papes_Interactive <- 
  ChangeWorkflow(workflow = Feng_Papes,
                   output = Chain(PerformanceMeasures, 
                                  PrintMap, 
                                  InteractiveMap, 
                                  ResponseCurve))
```

![The interactive map module overlays the raw occurrence data and predicted distribution over a global map, allowing users to interactively explore their results.\label{fig:interactive}](./figs/interactive_map.png)

>>>>

---
title: "The `zoon` R package for reproducible and shareable species distribution modelling"
author: "Nick Golding, Tom August, Tim C.D. Lucas, Emiel Van Loon & Greg McInerny"
bibliography: zoon_app.bib
csl: mee.csl
output:
  pdf_document:
    fig_caption: yes
    includes:  
      in_header: latex_preamble.tex
---

@@

# plot the saved image in the first panel
library(raster)
r <- brick('figs/workflow_basic.png')
par(mar = c(1, 1, 2, 1))
plotRGB(r, maxpixels = Inf)

@@

```{r fengworkflow, eval = TRUE, dpi = 300, fig.show = 'hide', results = 'hide', message = FALSE}
Feng_Papes <- 
  workflow(occurrence = SpOcc('Dasypus novemcinctus',
                              extent = c(-130, -20, -60, 60)),
            covariate = Bioclim(extent = c(-130, -20, -60, 60),
                                layers = c(1:4, 6, 9, 10, 12, 15)),
              process = Chain(Clean, 
                              MESSMask,  
                              Background(n = 10000, bias = 200), 
                              Crossvalidate(k = 5)),
                model = MaxEnt,
               output = PrintMap)

@@

![Map of the presence data (red), background data (black) and the MaxEnt relative probability of occurrence generated by the `PrintMap` module in the workflow `Feng_Papes`. White areas are masked due to being in the sea or outside the MESS mask. \label{fig:fengmap}](./figs/fengworkflow-1.png)

@@

```{r fengChangeWorkflow, eval = TRUE, fig.show = 'hide'}
Feng_Papes_Update <- 
  ChangeWorkflow(workflow = Feng_Papes,
                   output = Chain(InteractiveMap, 
                                  ResponseCurve(cov = 1)))
```

```{r combinefengupdate, eval = TRUE, echo = FALSE, dpi = 300, fig.show = 'hide', fig.height = 3.2}
# combine the printed map and first effect plot

# load both images
r_map <- brick('figs/interactive_map.png')
r_resp <- brick('figs/fengChangeWorkflow-2.png')

# set up layout so that heights are the same for both panels
plot_height <- min(nrow(r_map), nrow(r_resp))

# get rescaling factor for each
rescale_map <- plot_height / nrow(r_map)
rescale_resp <- plot_height / nrow(r_resp)

# get widths
width_map <- ncol(r_map) * rescale_map
width_resp <- ncol(r_resp) * rescale_resp

# set up layout (maximum total of 200 columns) with a gap in between
# mar won't work for these
gap <- 2
widths <- c(width_map, width_resp)
widths <- round(widths * ((200-gap) / sum(widths)))
widths <- c(widths[1], gap, widths[2])
mat <- matrix(rep(1:3, widths),
              nrow = 1)
layout(mat)

# plot interactive map
plotRGB(r_map, maxpixels = Inf)

# add a panel letter
mtext(text = 'A',
        side = 3,
        line = -3,
        adj = 0)

# gap
plot.new()

# plot the response curve
plotRGB(r_resp, maxpixels = Inf)

# add a panel letter
mtext(text = 'B',
        side = 3,
        line = -3,
        adj = 0)

![The outputs of the updated workflow object `Feng_Papes_Update`. A) A screenshot of the interactive map produced by the `InteractiveMap`, displaying raw occurrence data and predicted distribution over a global map, allowing users to interactively explore their results. B) A response curve produced by the `ResponseCurve` for the first covariate, bio1. \label{fig:interactive}](./figs/combinefengupdate-1.png)
